Is this normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-07-06 paper.qmd
added orcid id

---
title: "Is this normal? A new projection pursuit index to assess a sample against a multivariate null distribution"
author: 
  - name: Annalisa Calvi
    affiliations:
      name: Monash University
      department: School of Mathematics
  - name: Ursula Laa
    affiliations:
      name: University of Natural Resources and Life Sciences, Vienna
      department: Institute of Statistics
    email: ursula.laa@boku.ac.at
    ORCID: 0000-0002-0249-6439
  - name: Dianne Cook
    affiliations:
      name: Monash University
      department: Department of Econometrics and Business Statistics
    email: dicook@monash.edu
    ORCID: 0000-0002-3813-7155
format:
  jasa-pdf:
    keep-tex: true    
abstract: |
  Many data problems contain some reference or normal conditions, upon which to compare newly collected data. This scenario occurs in data collected as part of clinical trials to detect adverse events, and or for measuring climate change against historical norms. The data is typically multivariate, and often the normal ranges are specified by a multivariate normal distribution. The work presented in this paper develops methods to compare the new sample against the reference distribution with high-dimensional visualisation. It uses a projection pursuit guided tour to produce a sequence of low-dimensional projections steered towards those where the new sample is most different from the reference. A new projection pursuit index is defined, and the drawing of the projected ellipse is computed analytically. The methods are implemented in the R package, `tourr`. 
keywords:
  - exploratory data analysis
  - statistical graphics
  - dimension reduction
  - high-dimensional visualisation
  - climate change
  - clinical trials
bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{amsthm}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
  \newcommand\bD{$1\text{-}D$}
  \newtheorem*{theorem}{Theorem}
---

>>>>

---
title: "Is this normal? A new projection pursuit index to assess a sample against a multivariate null distribution"
author: 
  - name: Annalisa Calvi
    affiliations:
      name: Monash University
      department: School of Mathematics
    email: annalisa.calvi@monash.edu
    ORCID: 0009-0000-2977-0074
  - name: Ursula Laa
    affiliations:
      name: University of Natural Resources and Life Sciences, Vienna
      department: Institute of Statistics
    email: ursula.laa@boku.ac.at
    ORCID: 0000-0002-0249-6439
  - name: Dianne Cook
    affiliations:
      name: Monash University
      department: Department of Econometrics and Business Statistics
    email: dicook@monash.edu
    ORCID: 0000-0002-3813-7155
format:
  jasa-pdf:
    keep-tex: true    
abstract: |
  Many data problems contain some reference or normal conditions, upon which to compare newly collected data. This scenario occurs in data collected as part of clinical trials to detect adverse events, and or for measuring climate change against historical norms. The data is typically multivariate, and often the normal ranges are specified by a multivariate normal distribution. The work presented in this paper develops methods to compare the new sample against the reference distribution with high-dimensional visualisation. It uses a projection pursuit guided tour to produce a sequence of low-dimensional projections steered towards those where the new sample is most different from the reference. A new projection pursuit index is defined, and the drawing of the projected ellipse is computed analytically. The methods are implemented in the R package, `tourr`. 
keywords:
  - exploratory data analysis
  - statistical graphics
  - dimension reduction
  - high-dimensional visualisation
  - climate change
  - clinical trials
bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{amsthm}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
  \newcommand\bD{$1\text{-}D$}
  \newtheorem*{theorem}{Theorem}
---
